{\plant}: A package for modelling forest trait ecology \& evolution

### 2015-05-29 ms.tex
Minor tweaks

At the same time, we expect the {\plant} model to be useful for
modelling other phenomena. The internal structure of the {\plant}
model is similar in approach to the ``Ecosystem Demography''
\citep{Moorcroft-2001} and \textsc{lpj-guess} \citep{Smith-2014}
models. Like {\plant}, these models include size- structured
competition within a metapopulation of ``patches''. However, whereas
these models were designed to capture large-scale regional
biogeochemistry driven by climatic variables and rely on simulations,
{\plant} was designed so that the rules driving the dynamics of the
system could be manipulated and emergent demographic behaviours
studied deterministically and in detail. In {\plant} it is possible to grow plants in
isolation, in patches of competing plants and in disturbance-driven
metapopulations. Moreover, we put considerable effort into resolving
competitive hierarchies. 
The package is thus designed to sit in the middle
ground between a large-scale coupled simulation model and simple
abstract theoretical device.

Below we describe the general approach of the package, then provide
sections with details at nested levels of ecological organisation.
Detailed technical documents are provided as supplementary information
(see Appendices), updated versions of these technical documents will also
be made available within the {\plant} package itself.

@@

\section{Effects of size, trait and environment on demography of individual plants}

@@

Beyond its conceptual appeal, the idea of a dynamic equilibrium captured by
the trait-, size- and patch structured metapopulation model potentially
resolves several challenges faced by unstructured models. First, by scaling up
from patch to metapopulation, we close the demographic feedback loop to create
a self-regulating system. Unlike some vegetation models, there is no need to
set the density of seeds or seedlings, these arise as outputs from the model,
based on the combined effects of the model's physiological rules, species
traits, competitive interactions and the prevailing disturbance regime.

@@

The {\plant} package makes it easier for researchers and forest practitioners
alike to investigate a variety of non-linear demographic phenomena.  Rather
than provide a model with a few entry points, we have developed an extensible
framework that we hope can be used to answer a number of questions at a
variety of scales. The specific use-cases highlighted here arise from our
interest in the assembly of trait mixtures in communities. However a variety
of other uses would be possible and also extensions to address other aspects
of community assembly. As examples, we expect the package will prove useful
for investigating how morphological and physiological traits influence growth
rate and shade tolerance, and emergent phenomena such as multi-species self-
thinning, successional transitions and age-related changes in vegetation
productivity. The effects of competition
for multiple resources could likewise be investigated.

It is hoped that the approach here can be used to inform both future empirical
research and model development in neighbouring areas. As noted above, most
large-scale mechanistic models of plant growth do not explicitly model size
structure, so remain unable to investigate demographic phenomena
\citep{Cramer-2001, Sitch-2003, Dekauwe-2014}. At the same time,
much attention is paid to detailed leaf-level physiology. {\plant} aims to bridge
this gap that emerged in the model landscape, where detailed physiology and
size-structured dynamics combine to drive forest structure and function via
traits.

>>>>

At the same time, we expect the {\plant} model to be useful for
modelling other phenomena. The internal structure of the {\plant}
model is similar in approach to the ``Ecosystem Demography''
\citep{Moorcroft-2001} and \textsc{lpj-guess} \citep{Smith-2014}
models. Like {\plant}, these models include size- structured
competition within a metapopulation of ``patches''. However, whereas
these models were designed to capture large-scale regional
biogeochemistry driven by climatic variables and rely on simulations,
{\plant} was designed so that the rules driving the dynamics of the
system could be manipulated and emergent demographic behaviours
studied deterministically and in detail. In {\plant} it is possible to grow plants in
isolation, in patches of competing plants and in disturbance-driven
metapopulations. 
% TODO: can you change this?  the amount of effort we put in is
% decoupled from how useful other people may find it.
Moreover, we put considerable effort into resolving
competitive hierarchies. 
The package is thus designed to sit in the middle
ground between a large-scale coupled simulation model and simple
abstract theoretical device.

Below we describe the general approach of the package, then provide
sections with details at nested levels of ecological organisation.
Detailed technical documents are provided as supplementary information
(see Appendices), updated versions of these technical documents are
available within the {\plant} package itself.

@@

\section{Effects of size, trait and environment on demography of individual plants}Q

@@

The dynamic equilibrium captured by the trait-, size- and patch
structured metapopulation model potentially resolves several
challenges faced by unstructured models. First, by scaling up from
patch to metapopulation, we close the demographic feedback loop to
create a self-regulating system. Unlike some vegetation models, there
is no need to set the density of seeds or seedlings, these arise as
outputs from the model, based on the combined effects of the model's
physiological rules, species traits, competitive interactions and the
prevailing disturbance regime.

@@

The {\plant} package makes it easier for researchers and forest practitioners
alike to investigate a variety of non-linear demographic phenomena.  Rather
than provide a model with a few entry points, we have developed an extensible
framework that we hope can be used to answer a number of questions at a
variety of scales. The specific use-cases highlighted here arise from our
interest in the assembly of trait mixtures in communities. However a variety
of other uses would be possible and also extensions to address other aspects
of community assembly. As examples, we expect the package will prove useful
for investigating how morphological and physiological traits influence growth
rate and shade tolerance, and emergent phenomena such as multi-species self-
thinning, successional transitions and age-related changes in vegetation
productivity.

% TODO: In favour of dropping this whole paragraph?
It is hoped that the approach here can be used to inform both future
empirical research and model development in neighbouring areas.
{\plant} aims to bridge the gap between detailed physiological models
and large-scale mechanistic models of plant growth \citep{Cramer-2001,
  Sitch-2003, Dekauwe-2014}.  As noted above, most large-scale
mechanistic models of plant growth do not explicitly model size
structure, so remain unable to investigate demographic phenomena }. At
the same time, much attention is paid to detailed leaf-level
physiology. {\plant} aims to bridge this gap that emerged in the model
landscape, where detailed physiology and size-structured dynamics
combine to drive forest structure and function via traits.
